---
title: "Chemical lists and bulk workflows"
description: >
  Discover lists, resolve identifiers, preserve failures, and use bulk endpoints.
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

Bulk work is most reliable when you keep the original identifier and use the
DTXSID as the join key after resolution.

## Discover and retrieve CCD lists

```{r}
library(ComptoxR)
library(dplyr)

lists <- ct_chemical_list_all()
lists |>
  filter(grepl("fungic", paste(listName, label, shortDescription),
               ignore.case = TRUE)) |>
  select(listName, label, type, chemicalCount)

named_list <- ct_chemical_list_search_by_name("S11_SWISSPEST")
members <- ct_chemical_list_chemicals_search_by_listname("S11_SWISSPEST")
members |> select(any_of(c("dtxsid", "preferredName", "casrn")))
```

`ct_chemical_list_search_by_name()` uses a Dashboard list name. Use the full
list table for text discovery, then pass a selected `listName` to the two
specific functions.

## Resolve names and CAS numbers without losing failures

```{r}
input <- tibble(
  input_id = seq_len(4),
  identifier = c("benzene", "108-88-3", "not-a-real-chemical", "benzene")
)

resolved <- ct_chemical_search_equal_bulk(input$identifier)

resolution <- input |>
  left_join(
    resolved,
    by = c("identifier" = "searchValue"),
    relationship = "many-to-many"
  ) |>
  mutate(
    match_status = case_when(
      is.na(dtxsid) ~ "unmatched",
      duplicated(input_id) | duplicated(input_id, fromLast = TRUE) ~ "ambiguous",
      TRUE ~ "matched"
    )
  )

resolution |>
  select(input_id, identifier, dtxsid, preferredName, casrn, match_status)
```

The input row number preserves duplicates. The left join preserves unmatched
identifiers. More than one result for one `input_id` is explicit ambiguity.
Do not join endpoint data by name or CAS number after this step.

## Use the available bulk wrappers

```{r}
ids <- resolution |>
  filter(match_status == "matched") |>
  distinct(dtxsid) |>
  pull()

if (!length(ids)) stop("No unambiguous DTXSIDs were resolved.")

details <- ct_chemical_detail_search_bulk(ids)
property_exp <- ct_chemical_property_experimental_search_bulk(ids)
property_pred <- ct_chemical_property_predicted_search_bulk(ids)
fate <- ct_chemical_fate_search_bulk(ids)
exposure <- ct_exposure_seem_general_search_bulk(ids)
hazard <- ct_hazard_toxval_search_bulk(ids)
cancer <- ct_hazard_cancer_search_bulk(ids)
genetox <- ct_hazard_genetox_search_bulk(ids)
bioactivity <- ct_bioactivity_data_search_bulk(ids)
```

Each result can be joined to `resolution` by `dtxsid`. Empty endpoint results
must remain valid empty tables; do not treat them as proof that the chemical
has no hazard or exposure.

## Reverse the workflow

Endpoint results can define a new chemical set.

```{r}
selected_ids <- hazard |>
  filter(toxvalTypeSuperCategory == "Reference Value", !is.na(toxvalNumeric)) |>
  distinct(dtxsid) |>
  pull()

selected_details <- ct_chemical_detail_search_bulk(selected_ids)
selected_exposure <- ct_exposure_seem_general_search_bulk(selected_ids)
```

## Iterate only when no bulk wrapper exists

IVIVE has no bulk wrapper. Property summaries, fate summaries, production
volume, and bioactivity summaries also require iteration.

```{r}
ivive <- lapply(ids, ct_hazard_adme_ivive_search) |>
  bind_rows()
```

EPI Suite also has no package-level bulk wrapper. Keep any loop small and
rate-conscious.

```{r}
epi_results <- lapply(
  c("71-43-2", "108-88-3"),
  function(cas) {
    Sys.sleep(1)
    epi_submit(cas = cas, modules = "logKow,biodegradationRate")
  }
)
```

Use `batch_limit()` only for endpoints that accept package-managed batches.
It does not turn a single-chemical endpoint into a bulk endpoint.
