---
title: "Process batches and inspect requests"
description: >
  Query multiple chemicals and troubleshoot requests safely.
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

ComptoxR accepts vectors of identifiers. Bulk functions and the package batch
settings help process larger lists without writing a request loop yourself.

## Start with a small vector

```{r}
library(ComptoxR)

ids <- c("DTXSID7020182", "DTXSID7020192")
results <- ct_chemical_detail_search_bulk(ids)
results
```

Use the bulk function for the service you need:

```{r}
hazards <- chemi_hazard_bulk(ids)
toxicity <- ct_hazard_cancer_search_bulk(ids)
```

## Check the batch limit

The package has a default batch limit for APIs that accept batched requests.
Inspect or change it for the current session:

```{r}
batch_limit()
batch_limit(100)
```

Use a limit supported by the service and consider the size of the returned
data when processing large lists.

## Inspect a request without sending it

Debug mode builds dry-run requests. This is useful for checking an endpoint or
payload before using a real API key:

```{r}
run_debug(TRUE)
ct_chemical_detail_search("DTXSID7020182")
run_debug(FALSE)
```

Verbose mode reports progress while a request is running:

```{r}
run_verbose(TRUE)
```
