---
title: "Building a CCD chemical profile"
description: >
  Build a task-oriented CompTox profile for one chemical.
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

This article builds one Chemical Dashboard profile for benzene. It uses
`DTXSID7020182` as the join key. Keep units, sources, and evidence categories
when you combine results.

## Start with chemical identity

```{r}
library(ComptoxR)
library(dplyr)

dtxsid <- "DTXSID7020182"
detail <- ct_chemical_detail_search(dtxsid)

glimpse(detail)
detail |>
  select(dtxsid, preferredName, casrn, molFormula, averageMass, smiles)
```

A verified production response has one row and includes `dtxsid`,
`preferredName`, `casrn`, `molFormula`, `averageMass`, and `smiles`, with
additional identifiers and Dashboard summary fields.

## Add physical-chemical properties

Property names are service values. Inspect them before you build a large
query.

```{r}
ct_chemical_property_experimental_name()
ct_chemical_property_predicted_name()

water_summary <- ct_chemical_property_summary_search(
  dtxsid,
  propName = "Water Solubility"
)
water_experimental <- ct_chemical_property_summary_experimental_search(
  dtxsid,
  propName = "Water Solubility"
)
water_predicted <- ct_chemical_property_summary_predicted_search(
  dtxsid,
  propName = "Water Solubility"
)

water_summary |>
  select(propName, unit, experimentalMedian, experimentalCount,
         predictedMedian, predictedCount)
water_experimental |> select(propValue, sourceName, publicSourceUrl)
water_predicted |> select(propValue, sourceName, qmrfUrl)
```

The summary combines counts and simple statistics. Experimental rows give
reported values and sources. Predicted rows give model values and, when
available, model-report links. Do not compare values until you confirm their
units and basis.

## Add environmental fate

The fate wrappers use the same `propName` pattern. Some chemicals have only
experimental or only predicted evidence.

```{r}
bcf_summary <- ct_chemical_fate_summary_search(
  dtxsid,
  propName = "Bioconcentration Factor"
)
bcf_experimental <- ct_chemical_fate_experimental_search(
  dtxsid,
  propName = "Bioconcentration Factor"
)
bcf_predicted <- ct_chemical_fate_predicted_search(
  dtxsid,
  propName = "Bioconcentration Factor"
)

bcf_summary |>
  select(propName, unit, experimentalMedian, experimentalCount,
         predictedMedian, predictedCount)
bcf_experimental |> select(propValue, sourceName, publicSourceUrl)
bcf_predicted |> select(propValue, sourceName, qmrfUrl)
```

Other current fate property names include `"Biodeg. Half-Life"`,
`"Atmos. Hydroxylation Rate"`, and `"Soil Adsorp. Coeff. (Koc)"`.

## Add hazard evidence

The CCD ToxVal endpoint is a fast public API path. Select a small set of fields
and keep the evidence metadata.

```{r}
toxval <- ct_hazard_toxval_search(dtxsid)

toxval |>
  filter(toxvalTypeSuperCategory %in% c("Point of Departure", "Reference Value")) |>
  select(dtxsid, source, toxvalType, toxvalTypeSuperCategory,
         toxvalNumeric, toxvalUnits, speciesCommon, studyType,
         quality, qcCategory)

genetox <- ct_hazard_genetox_search(dtxsid)
cancer <- ct_hazard_cancer_search(dtxsid)
toxref <- ct_hazard_toxref_data_search(dtxsid)
```

`toxvalNumeric` has meaning only with `toxvalUnits`. `toxvalType` states the
value type, while `toxvalTypeSuperCategory` groups evidence such as points of
departure and reference values. Source, species, study type, quality, and QC
fields help a reviewer judge the evidence.

## Retrieve IVIVE parameters

```{r}
ivive <- ct_hazard_adme_ivive_search(dtxsid)
ivive |>
  select(dtxsid, description, measured, predicted, unit, model,
         species, reference)
```

The current response separates measured and predicted values. Keep `unit`,
`model`, `species`, and `reference` with each parameter.

## Add exposure estimates

SEEM general exposure is the primary compact exposure result.

```{r}
seem <- ct_exposure_seem_general_search(dtxsid)
seem |>
  select(dtxsid, productionVolume, units, starts_with("probability"))

production <- ct_exposure_ccd_production_volume_search(dtxsid)
production |> select(dtxsid, name, amount)
```

The SEEM probabilities describe exposure-use categories. They are screening
estimates, not measured individual exposure. Production-volume records are a
short supporting line of evidence.

## Start compact, then inspect assay records

```{r}
bioactivity_summary <-
  ct_bioactivity_data_summary_search_by_dtxsid(dtxsid)
bioactivity_summary |>
  select(dtxsid, activeMc, totalMc, activeSc, totalSc, ntested, nhit,
         cytotoxMedianUm)

bioactivity_detail <- ct_bioactivity_data_search_by_dtxsid(dtxsid)
glimpse(bioactivity_detail)
```

The summary gives assay counts and cytotoxicity context before you work with
detailed assay records.

## Know which calls are bulk

Use bulk wrappers for chemical detail, experimental and predicted properties,
fate records, CCD ToxVal, cancer, genetox, SEEM general exposure, and detailed
bioactivity data. The property and fate summary wrappers, IVIVE, production
volume, and bioactivity summary wrappers require one DTXSID at a time. Use a
small, rate-conscious iteration only when no bulk wrapper exists. The next
article shows both patterns.
