---
title: "Cheminformatics search and AMOS"
description: >
  Search structures, run available models, cluster chemicals, and use AMOS classifications.
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

Cheminformatics searches have different input contracts and result shapes.
Start with one search type and inspect its result before you combine searches.

## Search chemical space

```{r}
library(ComptoxR)
library(dplyr)

exact <- chemi_search("DTXSID7020182", search_type = "exact")
similar <- chemi_search(
  "DTXSID7020182",
  search_type = "similar",
  min_similarity = 0.9
)
substructure <- chemi_search("c1ccccc1", search_type = "substructure")
mass <- chemi_search(
  search_type = "mass",
  mass_type = "mono",
  min_mass = 77,
  max_mass = 79
)
features <- chemi_search(
  search_type = "features",
  element_include = c("C", "H"),
  exclude_all_others = TRUE
)
```

Exact, similarity, and substructure searches require a structure or chemical
query. Mass search requires a mass type and range. Feature search uses element
and structure flags.

Hazard search uses service categories instead of a structure query.

```{r}
hazard <- chemi_search(
  search_type = "hazard",
  hazard_name = "cancer",
  min_toxicity = "H",
  min_authority = "auth"
)
```

## Predict and cluster

```{r}
prediction <- chemi_predictor_models_predict(
  model_id = 1065,
  identifier = "DTXSID7020182"
)

clusters <- chemi_resolver_getsimilaritymap(
  c("DTXSID7020182", "DTXSID0020319", "DTXSID9020243"),
  idType = "DTXSID",
  format = "cluster"
)
plot(clusters$hc)
```

Model IDs belong to models deployed by the service. A known ID can be removed,
replaced, or unavailable on the selected server. Record the model ID and
service environment with each prediction. Treat availability as
service-controlled, not as a package guarantee.

## Inspect AMOS analytical methods

AMOS method discovery returns analytical method records. It is not a chemical
lookup.

```{r}
methods <- chemi_amos_method_list()
glimpse(methods)
head(methods)
```

Do not use `chemi_amos_method_with_spectra()` for a compound lookup. It requires
an AMOS method or spectrum record ID.

## Move through ClassyFire categories

First get the four ClassyFire levels for benzene.

```{r}
classification <-
  chemi_amos_get_classification_for_dtxsid("DTXSID7020182") |>
  bind_rows()

classification |>
  select(any_of(c("kingdom", "superklass", "klass", "subklass")))
```

Then enumerate valid child categories. Use `klass`; `class` is not a supported
request parameter.

```{r}
classes <- chemi_amos_next_level_classification(
  kingdom = classification$kingdom[[1]],
  superklass = classification$superklass[[1]]
)

subclasses <- chemi_amos_next_level_classification(
  kingdom = classification$kingdom[[1]],
  superklass = classification$superklass[[1]],
  klass = classification$klass[[1]]
)
```

Move from one ClassyFire category back to matching substances.

```{r}
class_members <- chemi_amos_substances_for_classification(
  kingdom = classification$kingdom[[1]],
  superklass = classification$superklass[[1]],
  klass = classification$klass[[1]],
  subklass = classification$subklass[[1]]
)
```

Request ClassyFire fields for several DTXSIDs with the AMOS batch workflow.

```{r}
amos_batch <- chemi_amos_batch(
  dtxsids = c("DTXSID7020182", "DTXSID0020319", "DTXSID9020243"),
  include_classyfire = TRUE
)
glimpse(amos_batch)
```

ClassyFire describes chemical structure classes. ChemFuncT describes
functional uses. Keep the two classification systems separate. Advanced users
can query ChemFuncT with
`chemi_amos_functional_uses_for_dtxsid("DTXSID7020182")`.
