---
title: "Search for chemicals and retrieve profiles"
description: >
  Find chemicals with common identifiers and retrieve CompTox records.
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

Chemical work usually starts with an identifier. ComptoxR accepts common
identifiers such as DTXSIDs, CAS numbers, names, SMILES, and InChI strings.

## Retrieve a chemical record

Use `ct_chemical_detail_search()` when you already have a DTXSID:

```{r}
library(ComptoxR)

benzene <- ct_chemical_detail_search("DTXSID7020182")
benzene
```

For several DTXSIDs, use the bulk function:

```{r}
benzene_and_toluene <- ct_chemical_detail_search_bulk(
  c("DTXSID7020182", "DTXSID7020192")
)
```

## Search by name or text

Use the chemical search functions when you do not know the DTXSID:

```{r}
ct_chemical_search_equal("benzene")
ct_chemical_search_contain("benzene", top = 20)
```

Use `chemi_resolver_lookup()` when you want identifier resolution across
several identifier types:

```{r}
chemi_resolver_lookup("benzene", idType = "name", fuzzy = TRUE)
```

## Inspect the result

The functions return tibbles by default. Use standard R tools to select or
save columns:

```{r}
benzene[c("dtxsid", "preferred_name")]
write.csv(benzene, "benzene.csv", row.names = FALSE)
```

Start with one chemical while learning the result shape. Move to bulk
functions after the single-chemical call gives the fields you need.
