---
title: "Development Lifecycle"
description: >
  How a change travels from a feature branch to a released version of
  ComptoxR, and how the GitHub Actions workflows build, test, version, and
  publish the package along the way.
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This article is for **contributors and maintainers**. It explains the branch
model, which GitHub Actions run at each step, and how a release bumps the
version and publishes the package. For per-workflow detail, read
[`.github/workflows/README.md`](https://github.com/seanthimons/ComptoxR/blob/main/.github/workflows/README.md);
this article ties those workflows together into one flow.

## The branch model

Feature branches feed two long-lived branches:

```
feature branch  ->  integration  ->  main  ->  Release (tag + GitHub release)
(feat/ fix/         (light            (trunk;      (manual, version-bumped)
 chore/ docs/)       staging)          full gate)
```

- **Feature branches** (`feat/`, `fix/`, `chore/`, `docs/`, ...) are cut from
  `integration`. All day-to-day work happens here.
- **`integration`** is light staging. Merges land here first so several
  features can settle together. This is a branch role, not an API server.
  Pushes run secret scanning and the integration package publisher. The
  publisher requires a `.9000` development version; stable metadata stops
  publication. It does not change the version itself.
- **`main`** is the trunk. The full check suite runs on every pull request that
  **targets `main`**, so `main` stays release-ready at all times.
- **Release** is a manual step run from `main`. It bumps the version, tags,
  regenerates `NEWS.md`, and cuts a GitHub release.

Branch and commit names follow Conventional Branch / Conventional Commits
(`type/description` and `type: description`). See `CONTRIBUTING.md`.

## Endpoint and generation rules

The public package ships operations from approved production schemas and the
local ECOTOX and ToxVal Plumber APIs. Staging and development server selectors,
fallback hooks, and operations without a production equivalent are removed.
The approved production EPI Suite URL remains `https://episuite.dev/api`.

Explicit service URLs remain supported. Resolution uses a package option,
then the corresponding environment variable, then the default URL:

```{r}
options(ComptoxR.chemi_burl = "https://your-server.example/api")
```

A URL override changes the target of existing wrappers. It does not add
operations that are absent from the public package. Loading the package does
not switch endpoint settings based on the date.

Install the checksum-pinned `specmill` toolkit from `dev/toolkit-lock.json`
before running development generation or validation:

```{r}
source("dev/install_toolkit.R")
install_toolkit()
```

Use Air 0.9.0 for generation, as CI does. The toolkit is a development
dependency; installed ComptoxR clients do not need it. Runtime hooks and
manual wrappers stay in ComptoxR. Follow the generation and validation
commands in
[CONTRIBUTING.md](https://github.com/seanthimons/ComptoxR/blob/main/CONTRIBUTING.md),
including `dev/check_public_api.R` before proposing a change.

Keep development and staging schemas and generated clients outside every
public ComptoxR checkout. For a separate local client, source
`dev/generate_local_client.R` and call `generate_local_client()` with explicit
input and output roots, a schema file, a base URL, and a distinct package name.
Promote an operation by regenerating it from its approved production schema.
Do not copy a local generated client into the public package.

## Which checks run, and when

The workflows are wired by branch and event, so the cost of CI scales with how
close a change is to release.

| Event | Workflows that run |
| --- | --- |
| Push to any branch (incl. `integration`) | Gitleaks Secret Scan |
| Push to `integration` | Publish Integration Package; requires a `.9000` version |
| Push to `dev-*` | Quick Tests |
| Push to `main` / `dev` | Test Coverage and Coverage Check (+ R-CMD-check on `main`) |
| **PR targeting `main`** | R-CMD-check, CRAN Readiness, Pipeline Integration Tests, Quick Tests, Test Coverage, Coverage Check, Commit Lint, documentation build |
| PR review **approved** | Rebase branches on PR approval (rebases open WIP branches) |

Path filters can skip code checks for documentation-only changes. The workflow
files define the exact triggers. The heavy R gate lives on the **PR into `main`**. `CRAN Readiness` is the
blocking one: it runs the CRAN-safe test lane (`COMPTOXR_CRAN_SAFE_TESTS=true`,
`NOT_CRAN=false`) with no API key, no live network, and no local database, so a
green PR means the package is CRAN-clean.

Tests run in three lanes — CRAN-safe unit/contract, replay/fixture, and
explicit live-recording. See `dev/TESTING_GUIDE.md` and
`.github/workflows/README.md` for the full lane rules.

## Scheduled automation

Several workflows run on a cron and open PRs or refresh data on their own:

| Workflow | Schedule (UTC) | What it does |
| --- | --- | --- |
| Schema Update Check | Mon/Wed/Fri 09:00 | Compares approved production schemas; opens a PR on `automated/schema-update` with regenerated wrappers, documentation, and tests |
| Build & Upload DSSTox Database | Mon 08:00 | Rebuilds and uploads the DSSTox database |
| Build & Upload Source-only ECOTOX Database | Mon 09:00 | Publishes native source data after checking that derived life-stage tables are absent |
| Build & Upload ToxValDB Database | Mon 10:00 | Rebuilds and uploads the ToxValDB database |
| API compatibility probe | Mon 12:00 | Tests direct and wrapper contracts at the effective configured URLs, which default to production |

Treat an `automated/schema-update` PR like any other change: review the diff,
let the `main` gate run, then merge.

Schema checks use the pinned toolkit and detect added, changed, and removed
schemas. They update documentation and tests for every schema change and
check the public boundary even when no schema changes occur. Failed downloads
preserve available production snapshots; they do not select a staging host.

The live API probe requires `CTX_API_KEY` and publishes a sanitized result
matrix. Known direct upstream defects remain visible. Wrapper contract
failures and configuration errors fail the job. The probe does not record
cassettes or cycle through staging and development hosts.

Life-stage mapping is an explicit, versioned envharmonizer step; it is not
part of database publication. Use a harmonizer release that supports the
database's recorded ECOTOX release. Keep local databases and mapping evidence,
update both client and server packages, and restart local Plumber servers.
See [ECOTOX](ecotox.html) for the upgrade steps. Do not rerun the retired
`db-ecotox.yml` workflow or publish a legacy derived database.

Before publishing a replacement that needs new harmonizer support, publish
and verify that experimental harmonizer release first. Verify the replacement
database and matching clients before withdrawing affected older downloads.

## Releasing

Releases are **manual and never hand-bumped**. Do not edit the `Version:` field
in `DESCRIPTION` yourself — the Release workflow owns it.

1. Land all intended changes on `main`.
2. GitHub -> Actions -> **Release** -> *Run workflow*, and choose
   `version_type`:
   - `patch` -> `1.5.0` to `1.5.1`
   - `minor` -> `1.5.0` to `1.6.0`
   - `major` -> `1.5.0` to `2.0.0`
   - `none`  -> release the current version as-is (no bump)

The workflow then, on `main`:

1. Runs `desc::desc_bump_version(<type>)` to bump `DESCRIPTION` (skipped for
   `none`).
2. **Guards** against re-releasing: if `v<version>` already exists as a tag, it
   aborts rather than overwrite.
3. Runs the package check and stops on warnings or errors.
4. Commits `chore: bump version to <version> [skip ci]` and tags `v<version>` locally.
5. Regenerates `NEWS.md` with `autonewsmd` (rendered via Quarto), commits
   `docs: update NEWS.md for v<version> [skip ci]`, and re-points the tag.
6. Builds the source archive and checks both source and archive against the
   public API boundary before publication.
7. Pushes the commits and tag, then runs `gh release create v<version>` with the
   source tarball attached.

Because `NEWS.md` is generated from commit history, the first line of each
commit message is the release-note text — write it as the note you want to
appear.

### Other build/publish workflows

These workflows do **not** create a stable release. All support manual
dispatch; the integration publisher also runs on pushes to `integration`:

- **Publish Rolling Package** — builds from `main` and moves the
  `package-latest` prerelease tag to that commit. A moving "latest dev build",
  separate from stable releases.
- **Build Package Artifact** — builds a source tarball and uploads it as a
  workflow artifact (retained 14 days). No release, no tag.
- **Publish Integration Package** — builds source, Windows, and macOS assets
  for `integration-latest` when the version ends in `.9000`. Each filename
  records the source commit. It does not publish development API clients.
- **Record VCR Cassettes** — the only sanctioned path for live cassette
  recording; requires the `CTX_API_KEY` secret. Ordinary CI never records from
  production.

The artifact and rolling package workflows check the public API boundary
before building. A prerelease package follows the same endpoint rules as a
stable release. Documentation builds check the rendered site as well.

## Putting it together

A typical change:

1. `git switch integration && git switch -c feat/my-change`
2. Commit with Conventional Commits; push. Gitleaks runs.
3. Open a PR into `main` (directly, or via `integration` first for staged
   work). The full gate runs; `CRAN Readiness` must be green.
4. Merge once checks pass.
5. When ready to ship, run the **Release** workflow from `main` with the right
   `version_type`.
