---
title: "ECOTOX"
description: >
  Install local ECOTOX data and run one focused ecological-effects query.
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

The ECOTOX workflow uses a local DuckDB database or a service that you host.
Install the data once, then inspect its health and schema.

## Install and inspect ECOTOX

```{r}
library(ComptoxR)
library(dplyr)

eco_install()
eco_health()
eco_tables()
eco_fields("results")
eco_inventory()
eco_species("rainbow trout", field = "common_name")
```

These calls can use a large local download, so site builds do not run them.

## Run a focused query

Filter by chemical, species, and endpoint in `eco_results()`. Results retain
the original ECOTOX life-stage code and description.

```{r}
trout <- eco_results(
  casrn = "71-43-2",
  common_name = "Rainbow trout",
  endpoint = "LC50"
) |>
  select(test_cas, common_name, latin_name, endpoint,
         organism_lifestage, org_lifestage, final_conc, obs_duration,
         final_obs_duration)
```

Keep concentration and duration units from the full result when you report
values. Apply life-stage harmonization explicitly with a versioned
`envharmonizer` release that matches the ECOTOX source release.

## Upgrade from derived life-stage output

This is a breaking change. Results no longer include `harmonized_life_stage`
or `reproductive_stage`. Remove the former `lifestage_details` argument from calls.
The `con` argument is now in position 11 instead of position 12; use its name
in calls that supply a connection.

Update the ComptoxR client and the package used by the local Plumber server.
Restart the server after the update. Existing local databases continue to
work. Queries ignore old derived tables and do not change or delete them.
Keep your local databases and mapping evidence for reproducibility.

## Advanced query design

Before a complex query, inspect `eco_fields()` for the current table names and
columns. Full ECOTOX query design can include effect groups, exposure types,
taxonomic joins, unit conversions, and study-level joins. Those larger joins
are outside this focused first workflow.
