---
title: "ECOTOX"
description: >
  Install local ECOTOX data and run one focused ecological-effects query.
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

The ECOTOX workflow uses a local DuckDB database or a service that you host.
Install the data once, then inspect its health and schema.

## Install and inspect ECOTOX

```{r}
library(ComptoxR)
library(dplyr)

eco_install()
eco_health()
eco_tables()
eco_fields("results")
eco_inventory()
eco_species("rainbow trout", field = "common_name")
```

These calls can use a large local download, so site builds do not run them.

## Run a focused query

Filter by chemical, species, and endpoint in `eco_results()`. Request detailed
life-stage fields, then filter life stage in R.

```{r}
trout <- eco_results(
  casrn = "71-43-2",
  common_name = "Rainbow trout",
  endpoint = "LC50",
  lifestage_details = TRUE
) |>
  filter(harmonized_life_stage == "Juvenile") |>
  select(test_cas, common_name, latin_name, endpoint,
         harmonized_life_stage, final_conc, obs_duration,
         final_obs_duration)
```

`eco_results()` does not have a direct life-stage filter argument. The
`lifestage_details = TRUE` result supplies harmonized fields for a clear
post-query filter. Keep concentration and duration units from the full result
when you report values.

## Advanced query design

Before a complex query, inspect `eco_fields()` for the current table names and
columns. Full ECOTOX query design can include effect groups, exposure types,
taxonomic joins, unit conversions, and study-level joins. Those larger joins
are outside this focused first workflow.
