---
title: "EPI Suite"
description: >
  Find a chemical, submit selected EPI Suite modules, and inspect nested estimates.
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

EPI Suite estimates physical-chemical, fate, and ecotoxicity properties. Search
first, then submit a selected chemical.

## Search and submit one compound

```{r}
library(ComptoxR)

hits <- epi_search("benzene")
str(hits, max.level = 1)

epi <- epi_submit(
  cas = "71-43-2",
  modules = paste(
    c("logKow", "biodegradationRate", "bioconcentration",
      "atmosphericHalfLife", "ecosar"),
    collapse = ","
  )
)
```

The result is nested by EPI module. Inspect names before selecting fields.

```{r}
names(epi$result)
selected <- epi$result[c(
  "logKow",
  "biodegradationRate",
  "bioconcentration",
  "atmosphericHalfLife",
  "ecosar"
)]
str(selected, max.level = 2)
```

Keep the module name, estimate, units, and any model warnings together. EPI
results are screening estimates, not measured values.

## Bulk limits

ComptoxR has no package-level bulk wrapper for EPI Suite. For two or three
chemicals, use the small, delayed loop in
[Chemical Lists and Bulk Workflows](batch-processing.html). Do not submit a
large vector as if `epi_submit()` were a bulk endpoint.
