---
title: "Getting started with ComptoxR"
description: >
  Install ComptoxR, configure an API key, and retrieve chemical data.
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

ComptoxR provides R functions for querying the U.S. EPA CompTox Chemical
Dashboard and related services. This guide assumes that you are new to R
packages or API-based data access.

## Install ComptoxR

Run this once in the R console:

```{r}
install.packages("pak")
pak::pkg_install("seanthimons/ComptoxR")
```

Load the package whenever you start a new R session:

```{r}
library(ComptoxR)
```

## Configure access

Some services require an API key. Request one by emailing
`ccte_api@epa.gov` with the subject `API Key Request`.

For a persistent setup, open your user `.Renviron` file:

```{r}
file.edit("~/.Renviron")
```

Add the following line, replacing the placeholder with your key:

```text
ctx_api_key=YOUR_KEY_HERE
```

Save the file and restart R. Then check the setup:

```{r}
ct_api_key()
run_setup()
```

Do not put an API key directly in a script that you share or commit.

## Retrieve a chemical record

Many functions accept a DTXSID, CAS number, chemical name, or another chemical
identifier. For example:

```{r}
result <- ct_chemical_detail_search("DTXSID7020182")
result
```

Several identifiers can be queried together:

```{r}
result <- ct_chemical_detail_search(
  c("DTXSID7020182", "DTXSID7020183")
)
result
```

The returned object is normally a tibble, which can be inspected, filtered,
and saved with familiar R tools.

## Find the right function

The [reference](../reference/index.html) lists all exported functions. The
function names use prefixes to show the service family:

- `ct_` functions query CompTox Chemical Dashboard services.
- `chemi_` functions query cheminformatics and chemical-service APIs.
- `dss_`, `eco_`, and `toxval_` functions work with local EPA databases.
- `epi_` and `genra_` functions access prediction services.

Start with a single chemical and a small result. Then move to batch functions
when the single-query version works for your data.

## Troubleshooting

Use verbose mode to see progress:

```{r}
run_verbose(TRUE)
```

Use debug mode to inspect a request without sending it:

```{r}
run_debug(TRUE)
```

Both settings apply only to the current R session.
