---
title: "Explore hazard and toxicity data"
description: >
  Retrieve hazard, safety, toxicity, and bioactivity information for chemicals.
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

ComptoxR connects a chemical identifier to several EPA hazard and toxicity
services. The results come from different sources, so keep the source fields
when saving or combining results.

## Hazard and safety information

For cheminformatics hazard and safety summaries:

```{r}
library(ComptoxR)

chemi_hazard("DTXSID7020182")
chemi_safety("DTXSID7020182")
```

For cancer and genotoxicity information from the CompTox Dashboard:

```{r}
ct_hazard_cancer_search("DTXSID7020182")
ct_hazard_genetox_search("DTXSID7020182")
```

## Toxicity and bioactivity

Query ToxValDB data for a DTXSID:

```{r}
toxval_search("DTXSID7020182")
```

Retrieve CompTox bioactivity data by DTXSID:

```{r}
ct_bioactivity_data_search_by_dtxsid("DTXSID7020182")
```

## Compare several chemicals

Many services provide bulk variants. Use them when you have a vector of
identifiers:

```{r}
chemi_hazard_bulk(c("DTXSID7020182", "DTXSID7020192"))
ct_hazard_cancer_search_bulk(c("DTXSID7020182", "DTXSID7020192"))
```

These results should be interpreted with the source database documentation;
the presence of a record does not by itself establish a risk conclusion.
