---
title: "Hazard data and ToxValDB"
description: >
  Use fast CCD hazard endpoints, then deepen the search with local ToxValDB data.
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

The CCD hazard endpoints are the fast API path. ToxValDB is a separate local
DuckDB database or a service that you host. It is not another public CCD
endpoint.

## Start with CCD hazard endpoints

```{r}
library(ComptoxR)
library(dplyr)

dtxsid <- "DTXSID7020182"

ccd_toxval <- ct_hazard_toxval_search(dtxsid)
ccd_toxval |>
  select(dtxsid, source, toxvalType, toxvalTypeSuperCategory,
         toxvalNumeric, toxvalUnits, speciesCommon, studyType,
         quality, qcCategory)

ct_hazard_genetox_search(dtxsid)
ct_hazard_cancer_search(dtxsid)
ct_hazard_toxref_data_search(dtxsid)
```

Use the CCD path for quick profile work and bulk screening. Keep source, value
type, units, species, and quality fields.

## Install and inspect ToxValDB

```{r}
toxval_install()
toxval_health()
toxval_tables()
toxval_fields("toxval")
head(toxval_sources())
```

`toxval_install()` downloads and configures the local database. These commands
can use a substantial download and are disabled during the site build. A
self-hosted Plumber service provides the same high-level query workflow.

## Search and filter local records

`toxval_search()` is the short DTXSID path.

```{r}
local_values <- toxval_search(dtxsid)
local_values |>
  select(dtxsid, source, toxval_type, toxval_numeric, toxval_units,
         species_common, qc_status, qc_category)
```

Use `toxval_results()` for focused filters.

```{r}
iris_oral <- toxval_results(
  dtxsid = dtxsid,
  source = "IRIS",
  toxval_type = "RfD",
  species = "Human",
  qc_status = "pass"
)

iris_oral |>
  select(dtxsid, source, toxval_type, toxval_numeric, toxval_units,
         species_common, study_type, qc_status, qc_category)
```

The filters accept vectors. `qc_status = "pass_or_not_determined"` is the
default. Use `"pass"` for accepted rows only or `"all"` when you must review
failed QC rows. Keep failed rows labeled; do not silently mix them with passed
records.

ToxValDB deepens the hazard review with local, filterable records. It does not
replace the identity, property, fate, exposure, or bioactivity parts of the
[CCD chemical profile](chemical-search.html).
