---
title: "Local data and development"
description: >
  Install local EPA data, inspect schemas, and find contributor guidance.
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

Local database copies support repeated queries, offline work, and controlled
data extraction. Downloads can be large. Install only the database that your
work needs.

## Install local databases

```{r}
library(ComptoxR)

dss_install()
eco_install()
toxval_install()
```

Use `source`, `tag`, or `overwrite` only when you need a specific release or
must replace an existing copy. See each function reference for its current
options.

## Check health and schema before querying

```{r}
eco_health()
eco_tables()
eco_fields("results")
eco_inventory()

toxval_health()
toxval_tables()
toxval_fields("toxval")
toxval_sources()
```

The health result identifies the configured local data source. Table and field
functions help you confirm the schema before a large query.

## Use the focused articles

See [Hazard Data and ToxValDB](hazard-and-toxicity.html) for local toxicity
values and [ECOTOX](ecotox.html) for ecological effects. DSSTox provides local
chemical search and resolution:

```{r}
dss_search("benzene", limit = 20)
dss_resolve("71-43-2")
```

## Develop ComptoxR

Read the standalone [development lifecycle](development-lifecycle.html) before
changing the package. It describes branches, checks, continuous integration,
and releases. Also read `CONTRIBUTING.md` in the repository root. Do not record
API cassettes unless the task requires production traffic and you can run the
cassette safety checks.
