---
title: "Use the local EPA databases"
description: >
  Install and query local DSSTox, ECOTOX, and ToxValDB copies.
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

Some ComptoxR services use local database files. Local copies are useful when
you need repeated queries, offline work, or more control over data extraction.
They can require substantial disk space and may take time to download.

## Install a database

Install the database you need. These commands may download a large file:

```{r}
library(ComptoxR)

dss_install()
eco_install()
toxval_install()
```

Use the `source`, `tag`, or `overwrite` arguments when you need a specific
release or want to replace an existing local copy. Check the function reference
for the current database options.

## Query DSSTox

Search the local DSSTox database by text:

```{r}
dss_search("benzene", limit = 20)
```

Resolve a CAS number:

```{r}
dss_resolve("71-43-2")
```

## Query ToxValDB

Use a DTXSID to retrieve local toxicity values:

```{r}
toxval_search("DTXSID7020182")
```

## Query ECOTOX

ECOTOX queries can filter by chemical, species, endpoint, and other fields:

```{r}
eco_results(casrn = "71-43-2", endpoint = "mortality")
```

Use `eco_fields()` and `eco_tables()` to inspect the available local schema
before writing a larger query.
