---
title: "Risk evidence compilation"
description: >
  Compile screening-level identity, fate, hazard, exposure, and bioactivity evidence.
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

This workflow compiles evidence that can support risk characterization or
screening. It is not a completed regulatory risk assessment. It does not
define an exposure scenario, select critical studies, derive points of
departure, or make a regulatory conclusion.

## Discover relevant CCD lists

```{r}
library(ComptoxR)
library(dplyr)

lists <- ct_chemical_list_all()
fungicide_lists <- lists |>
  filter(grepl(
    "fungic",
    paste(listName, label, shortDescription, longDescription),
    ignore.case = TRUE
  )) |>
  select(listName, label, type, chemicalCount, shortDescription)

fungicide_lists
ct_chemical_list_search_by_name("S11_SWISSPEST")
```

Dashboard lists can change. Use list discovery to find current candidates, but
use a small explicit input vector for a reproducible example.

## Resolve a fixed fungicide input

```{r}
input <- tibble(
  input_id = seq_len(4),
  identifier = c("captan", "chlorothalonil", "azoxystrobin",
                 "not-a-real-chemical")
)

resolved <- ct_chemical_search_equal_bulk(input$identifier)
resolution <- input |>
  left_join(
    resolved,
    by = c("identifier" = "searchValue"),
    relationship = "many-to-many"
  ) |>
  mutate(
    match_status = case_when(
      is.na(dtxsid) ~ "unmatched",
      duplicated(input_id) | duplicated(input_id, fromLast = TRUE) ~ "ambiguous",
      TRUE ~ "matched"
    )
  )

resolution |>
  select(input_id, identifier, dtxsid, preferredName, casrn, match_status)

ids <- resolution |>
  filter(match_status == "matched") |>
  distinct(dtxsid) |>
  pull()

if (!length(ids)) stop("No unambiguous DTXSIDs were resolved.")
```

Keep the failed input in `resolution`. Excluding it without a record would
make the evidence table look more complete than the input process was.

## Compile CCD evidence

Use bulk wrappers when they exist.

```{r}
identity <- ct_chemical_detail_search_bulk(ids)
property_experimental <- ct_chemical_property_experimental_search_bulk(ids)
property_predicted <- ct_chemical_property_predicted_search_bulk(ids)
fate <- ct_chemical_fate_search_bulk(ids)
hazard <- ct_hazard_toxval_search_bulk(ids) |>
  filter(toxvalTypeSuperCategory %in% c(
    "Point of Departure",
    "Reference Value"
  )) |>
  select(dtxsid, source, toxvalType, toxvalTypeSuperCategory,
         toxvalNumeric, toxvalUnits, speciesCommon, studyType,
         quality, qcCategory)
exposure <- ct_exposure_seem_general_search_bulk(ids)
bioactivity_detail <- ct_bioactivity_data_search_bulk(ids)
```

Use small iteration for endpoints without bulk wrappers.

```{r}
ivive <- lapply(ids, ct_hazard_adme_ivive_search) |>
  bind_rows()

bioactivity_summary <- lapply(
  ids,
  ct_bioactivity_data_summary_search_by_dtxsid
) |>
  bind_rows()
```

The result now contains chemical identity, physical-chemical evidence, fate
and transport, selected hazard evidence, IVIVE, SEEM exposure estimates, and
bioactivity summaries. DTXSID is the join key for every CCD table.

## Add optional local evidence

Run these calls only after you install the local databases.

```{r}
toxval_local <- toxval_results(
  dtxsid = ids,
  qc_status = "pass_or_not_determined"
)

cas_map <- identity |>
  distinct(dtxsid, casrn)

ecotox_local <- eco_results(
  casrn = cas_map$casrn,
  eco_group = "Fish",
  endpoint = c("LC50", "EC50"),
  lifestage_details = TRUE
) |>
  left_join(cas_map, by = c("test_cas" = "casrn"))
```

ToxValDB joins directly by DTXSID. ECOTOX starts with CASRN, so use the
identity table to add the DTXSID immediately. Keep unmatched CAS numbers
visible.

## Build an evidence-availability table

```{r}
evidence <- list(
  identity = identity,
  property_experimental = property_experimental,
  property_predicted = property_predicted,
  fate = fate,
  hazard = hazard,
  ivive = ivive,
  exposure = exposure,
  bioactivity_summary = bioactivity_summary
)

available <- bind_rows(lapply(names(evidence), function(evidence_type) {
  endpoint_result <- evidence[[evidence_type]]
  if (!"dtxsid" %in% names(endpoint_result)) {
    return(tibble(
      dtxsid = character(),
      evidence_type = character(),
      available = logical()
    ))
  }

  endpoint_result |>
    filter(!is.na(dtxsid)) |>
    distinct(dtxsid) |>
    mutate(evidence_type = evidence_type, available = TRUE)
}))

availability_by_dtxsid <- tidyr::crossing(
  dtxsid = ids,
  evidence_type = names(evidence)
) |>
  left_join(available, by = c("dtxsid", "evidence_type")) |>
  mutate(available = coalesce(available, FALSE)) |>
  tidyr::pivot_wider(
    names_from = evidence_type,
    values_from = available,
    values_fill = FALSE
  )

availability <- resolution |>
  distinct(input_id, identifier, dtxsid, match_status) |>
  left_join(availability_by_dtxsid, by = "dtxsid") |>
  mutate(across(all_of(names(evidence)), ~ coalesce(.x, FALSE)))

availability
```

An empty endpoint result means “not retrieved from this endpoint,” not “no
evidence exists.” Preserve that distinction in the screening summary.

```{r}
screening_summary <- identity |>
  select(dtxsid, preferredName, casrn, molFormula, averageMass) |>
  left_join(
    exposure |>
      select(dtxsid, productionVolume, units, starts_with("probability")),
    by = "dtxsid"
  ) |>
  left_join(
    bioactivity_summary |>
      select(dtxsid, ntested, nhit, cytotoxMedianUm),
    by = "dtxsid"
  )

screening_summary
```

Review the full source tables before interpretation. The compact table is a
screening index into the evidence, not a replacement for study review or a
regulatory risk assessment.
