---
title: "ComptoxR"
subtitle: "An R Package to Retrieve Compound Information from US EPA Products and APIs for Rapid Chemical Hazard, Risk, Toxicological Evaluation and Screening"
format: revealjs
author: Sean Thimons, ORISE-CESER/WID/BCB
editor: visual
---

```{r}
#| label: load-packages
#| include: false
#| eval: true
#| echo: false

library(ComptoxR)
library(dplyr)
library(stringr)
library(knitr)
library(kableExtra)
library(flextable)
library(tidyr)
library(purrr)

comptox_server(server = 1)

prod_volume <- readRDS('prod_volume.Rds')

```

# Introduction

## Why was it made

- Need for large, tailored data requests

- Bulk data requests from CCD weren't available or took a long time

- Needed a solution for working with large datasets; Excel spreadsheets wouldn't work.

## Implementation and Design Details

- Most functions that work to retrieve data on a compound are also implemented in a 'batch' function to reduce query times. 

- Typically, any function will use a RESTful GET request for a single compound, and a POST request for everything else.

- Most functions require API token from EPA (no cost!)

--------

## Implementation and Design Details

- When possible, the data returned is returned as-is from the API. 

- Some functions (e.g.: `ct_details()` offer different subsets of data from the endpoint. Where this is an option, the default option is specified to the user, with optional arguments detailed in the appropriate help file.

# Use cases

## Name searching

- Searching by string

```{r}
#| label: name-search1
#| eval: true
#| include: true
#| echo: true
#| output-location: slide

#String search
ct_name('atraz', param = 'start-with') %>%
  select(searchName, rank, preferredName) %>%
  head(n = 10)
```

- Seaching by InChlKey search

```{r}
#| eval: true
#| include: true
#| label: name-search2
#| echo: true
#| output-location: slide

#InChlKey search
ct_name('MXWJVTOOROXGIU', param = 'start-with') %>%
  select(searchValue, rank, preferredName) %>%
  head(n = 10)
```

- Seach by CAS-RN
```{r}
#| label: name-search3
#| eval: true
#| include: true
#| echo: true
#| output-location: slide

#CAS_RN search
ct_name('1912-24', param = 'contain') %>%
  select(searchValue, preferredName, casrn)
```

## Synonymn searching

```{r}
#| label: syno-search
#| eval: true
#| include: true
#| echo: true
#| output-location: slide

ct_synonym(query = 'DTXSID7020182') %>%
  head(n = 10)
```

## Compound details

- Seaching for compound details

```{r}
#| label: detail-search
#| eval: true
#| include: true
#| echo: true
#| output-location: slide

df <- ct_details(dtx_list[1])

# A selection of the data available for the query

df %>%
  select(dtxsid,preferredName, smiles, molFormula, monoisotopicMass,toxcastSelect,octanolWaterPartition,qcLevelDesc) %>%
  glimpse()
```

## Querying for physio-chemical properties 

### ...for Perfluorooctanoic acid

```{r}
#| label: prop_search
#| eval: true
#| include: true
#| echo: true
df <- ct_prop(query = dtx_list[1], debug = FALSE)

```

----

```{r}
#| label: prop_table
#| eval: true
#| echo: false
#| output-location: slide
df %>%
  select(name, value, source, propType, unit, dtxsid) %>% 
  group_by(name, propType) %>% 
  reframe(value = round(mean(value), 3), unit) %>%
  distinct(name, propType, value, unit) %>% 
  pivot_wider(., id_cols = c(name, unit), names_from = propType, values_from = value)
```

## ...and environmental fate & transport data

```{r}
#| label: eft-search
#| eval: true
#| include: true
#| echo: true

df <- ct_env_fate(query = dtx_list[1], debug = FALSE)

```
----
```{r}
#| label: eft_table
#| eval: true
#| echo: false
#| output-location: slide

df %>% 
  select(valueType, modelSource, unit, endpointName, resultValue) %>%
  group_by(endpointName, valueType) %>%
  summarize(value = round(mean(resultValue), 3), unit) %>%
  distinct(endpointName, valueType, value, unit) %>% 
  pivot_wider(., id_cols = c(endpointName, unit), names_from = valueType, values_from = value) 
```

## Non-Targeted Analysis Compound Searching
    
To better assist in the hazard and risk management space, several functions are suggested for use. 

- Searching by mass range: 

```{r}
#| label: nta-search1
#| eval: true
#| echo: true
#| include: true
#| output-location: slide


#Searching by mass range
    ct_search_mass(start = 413.9, end = 414) %>%
      select(dtxsid, preferredName, casrn, molFormula)
```    

----
    
- Searching by formula: 
```{r}
#| label: nta-search2
#| eval: true
#| echo: true
#| include: true
#| output-location: slide

#Searching by molecular formula
    ct_search_formula(query = 'C10H19O6PS2') %>%
      select(dtxsid, preferredName, casrn, averageMass, sourcesCount, toxcastSelect, qcLevel) %>%
      arrange(desc(sourcesCount))
```

## Searching for hazard records

```{r}
#| label: hazard
#| eval: true
#| echo: true

df <- ct_hazard(query = 'DTXSID4020375')


```
```{r}
#| label: hazard2
#| eval: true
#| echo: false
#| output-location: slide

df %>% group_by(riskAssessmentClass, supercategory, toxvalUnits) %>% summarize(toxvalNumeric = mean(toxvalNumeric))
```


## Searching for presence on aggregate lists

- Three functions for aggregate lists
  + Retrieve all public lists
  + Retrieve all compounds on a given list
  + Retrieve all public lists a compound is on

```{r}
#| label: lists_all
#| echo: true
#| output-location: slide
ct_lists_all() %>% select(name, type, label, longDescription)
```
----
```{r}
#| label: list_search1
#| echo: true
#| output-location: slide

ct_list('PRODWATER') %>%
  select(preferredName, casrn, smiles)
```

# Other functions of interest

## Other functions

- Hazard queries for GHS, skin/eye, genotoxicity, cancer records
- Queries for TEST and OPERA QSAR models
- Queries for [GenRA Generalized Read-Across]
- Hazard comparison table from Cheminformatics

# Remaining work

## Remaining work

- Bioactivity, in-vitro records not currently implemented
- Better error handling
- Compound structure picture retrieval
